---
title: "STAT 405/605 Presentation Group 11"
subtitle: "Insights into arXiv Articles Data"
author: "Gabe Gress, Minyue Hu, Tina Dantono, Kristin West"
date: "11/17/2020"
output:
  beamer_presentation: default
  ioslides_presentation: default
---
# Introduction

## Dataset

* Primary dataset: the arXiv Bulk Data from [the arXiv official website](https://arxiv.org/help/bulk_data)
  * arXiv, hosted by Cornell University, is a web-based open-access archive for over 1.7M scholarly articles
  * The documents in arXiv range from 1991-07 to 2020-10

* Secondary dataset: Universities & Colleges Dataset
  * Published by [the Department of Homeland Security](https://catalog.data.gov/dataset/colleges-and-universities-cb8a4), so the reliability is extremely high

## Background

- arXiv collects research papers on the topics of physics, mathematics, computer science, quantitative biology, quantitative finance, statistics, electrical engineering and systems science, and economics.

- As of October 2016, the submission rate has exceeded 10,000 articles per month.

- We chose to study the literature information on the arXiv website because the existence of this database is one of the factors that created the open access movement in the scientific publishing industry.

## Why?

- arXiv accepts publications free of charge, meaning it is most likely to fairly represent all scientific contributions
- Most preprints are on arXiv, so it often can represent trends in science

## Goals: 
- Our research debated whether different authors have different propensities for research topics over time

- We expected the final results to show that certain research fields are of greater interest over time

## Goals (cont.) :
- We also considered the following issues: 
  - which year has the greatest number of published papers on average
  - the number of papers published in each arXiv category per year
  - the geographical distribution of the publishers and the types of papers they published
  - the universities to which the researchers belong,
  - the influencing factors on the papers (i.e. funding)
  - 6 categories: physics, computer science, math, statistics, economics, electrical engineering and systems science. Categories can also be combined. 



# Data Exploration

## Growth of arXiv
```{r,echo=FALSE}
library(stringr)
library(ggplot2)
library(ggpubr)

arxiv <- read.csv(file="arxiv-dataset.csv",stringsAsFactors = FALSE)

yearsarx <- sapply(arxiv$id,as.character)
yearsarx <- str_extract(yearsarx, pattern="(\\d{4})")
datesarx <- as.Date(paste(yearsarx, "01",sep=""), "%y%m%d")

#Total Amount of ArXiv Paper over Time
ggplot(as.data.frame(yearsarx)) + 
  stat_count() + 
  aes(x = datesarx) +
  theme(panel.background = element_rect(fill = "white", colour = "black",
                                    size = 2, linetype = "solid"),
        panel.grid.major = element_line(size = 0.25, linetype = 'solid',
                                    colour = "black"),
        panel.grid.minor = element_line(size = 0.25, linetype = 'solid',
                                    colour = "black")) + 
  theme(plot.background = element_rect(fill = "white"))  +
  labs(title="ArXiv Papers Amount over Time", x="Time", y="Amount Count")

remove(yearsarx)

```

## Trends in Format
```{r,echo=FALSE}
#pages
commentsarx <- arxiv$comments
pagesarx <- str_extract(commentsarx, "(\\d)+(?= pages)")

#omit NA rows
newpagesarx <- as.data.frame(pagesarx)
Time_Pages <- data.frame("Time" = datesarx,
                         "Pages" = as.numeric(unlist(newpagesarx)))

Time_Pages <- Time_Pages[!is.na(Time_Pages$Pages), ]
Time_Pages$Time = (as.numeric(format(Time_Pages$Time,"%Y")))%/%5*5

ggplot(data = Time_Pages) +
  stat_summary(mapping = aes(x = Time, y = Pages),
               fun.min = function(z) { quantile(z,0.25) },
               fun.max = function(z) { quantile(z,0.75) },
               fun = mean, size = 1.3) +
  ggtitle("Relationship Between the Number of Arxiv Pages and Time") +
  theme(panel.background = element_rect(fill = "yellowgreen", colour = "black",
                                    size = 2, linetype = "solid"),
        panel.grid.major = element_line(size = 0.25, linetype = 'solid',
                                    colour = "black"),
        panel.grid.minor = element_line(size = 0.25, linetype = 'solid',
                                    colour = "black")) +
  theme(plot.background = element_rect(fill = "white"))

remove(commentsarx)
remove(pagesarx)
remove(newpagesarx)
remove(Time_Pages)
```


## Overall Distribution
```{r, echo=FALSE}
names(arxiv) <- c("X", "ID", "Submitter", "Authors", "Title", "Comments", 
                  "JournalRef", "DOI", "ReportNo", 
                  "Categories", "License", "Abstract", "Versions", 
                  "UpdateDate", "AuthorsParsed")
#Physics
arxiv$Categories[str_starts(arxiv$Categories, "astro-ph")] <- "Astrophysics"
arxiv$Categories[str_starts(arxiv$Categories, "cond-mat")] <- "Condensed Matter"
arxiv$Categories[str_starts(arxiv$Categories, "supr-con")] <- "Condensed Matter"
arxiv$Categories[str_starts(arxiv$Categories, "mtrl-th")] <- "Condensed Matter"
arxiv$Categories[str_starts(arxiv$Categories, "gr-qc")] <- "General Relativity and Quantum Cosmology"
arxiv$Categories[str_starts(arxiv$Categories, "hep")] <- "High Energy Physics"
arxiv$Categories[str_starts(arxiv$Categories, "math-ph")] <- "Mathematical Physics"
arxiv$Categories[str_starts(arxiv$Categories, "nlin")] <- "Nonlinear Sciences"
arxiv$Categories[str_starts(arxiv$Categories, "adap")] <- "Nonlinear Sciences"
arxiv$Categories[str_starts(arxiv$Categories, "chao")] <- "Nonlinear Sciences"
arxiv$Categories[str_starts(arxiv$Categories, "solv")] <- "Nonlinear Sciences"
arxiv$Categories[str_starts(arxiv$Categories, "patt-")] <- "Nonlinear Sciences"

arxiv$Categories[str_starts(arxiv$Categories, "nucl")] <- "Nuclear Physics"
arxiv$Categories[str_starts(arxiv$Categories, "physics")] <- "Other Physics"
arxiv$Categories[str_starts(arxiv$Categories, "acc")] <- "Other Physics"
arxiv$Categories[str_starts(arxiv$Categories, "atom")] <- "Other Physics"
arxiv$Categories[str_starts(arxiv$Categories, "ao")] <- "Other Physics"
arxiv$Categories[str_starts(arxiv$Categories, "bayes")] <- "Other Physics"
arxiv$Categories[str_starts(arxiv$Categories, "chem-")] <- "Other Physics"
arxiv$Categories[str_starts(arxiv$Categories, "plasm")] <- "Other Physics"

arxiv$Categories[str_starts(arxiv$Categories, "quant-ph")] <- "Quantum Physics"

#math
arxiv$Categories[str_starts(arxiv$Categories, "math")] <- "Math"
arxiv$Categories[str_starts(arxiv$Categories, "alg")] <- "Math"
arxiv$Categories[str_starts(arxiv$Categories, "q-")] <- "Math"
arxiv$Categories[str_starts(arxiv$Categories, "funct-")] <- "Math"
arxiv$Categories[str_starts(arxiv$Categories, "dg-ga")] <- "Math"

#Computer Science
arxiv$Categories[str_starts(arxiv$Categories, "CoRR")] <- "Computer Science"
arxiv$Categories[str_starts(arxiv$Categories, "cs")] <- "Computer Science"
arxiv$Categories[str_starts(arxiv$Categories, "cmp")] <- "Computer Science"
arxiv$Categories[str_starts(arxiv$Categories, "comp")] <- "Computer Science"

#the rest
arxiv$Categories[str_starts(arxiv$Categories, "q-bio")] <- "Quantitative Biology"
arxiv$Categories[str_starts(arxiv$Categories, "q-fin")] <- "Quantitative Finance"
arxiv$Categories[str_starts(arxiv$Categories, "stat")] <- "Statistics"
arxiv$Categories[str_starts(arxiv$Categories, "eess")] <- "Electrical Engineering and Systems Science"
arxiv$Categories[str_starts(arxiv$Categories, "econ")] <- "Economics"

category_arxiv <-  table(arxiv$Categories)

physics_tot <- as.numeric(category_arxiv)[1] + 
  as.numeric(category_arxiv)[3] +
  as.numeric(category_arxiv)[6] +
  as.numeric(category_arxiv)[7] + 
  as.numeric(category_arxiv)[9] + 
  as.numeric(category_arxiv)[10] +
  as.numeric(category_arxiv)[11] +
  as.numeric(category_arxiv)[12] +
  as.numeric(category_arxiv)[13]

category_arxiv_final <- data.frame(
  Category = c("Physics", "Math", 
              "Computer Science", 
              "Statistics", "Electrical Engineering and Systems Science",
              "Economics"), 
  Amount = c(physics_tot, as.numeric(category_arxiv)[8], 
             as.numeric(category_arxiv)[2], as.numeric(category_arxiv)[14],
             as.numeric(category_arxiv)[5], as.numeric(category_arxiv)[4]), 
  
  stringsAsFactors = FALSE
)

#plot for all subjects
plot1 <- ggplot(data = category_arxiv_final) + 
  aes(Category, Amount) +   
  geom_bar(stat='identity', width=0.5, color='red',fill='steelblue') +
  xlab("Categories") + 
  ylab("Number of Papers Submitted") + 
  ggtitle("Number of Papers Submitted by Category")  + 
  theme(panel.background = element_rect(fill = "#BFD5E3", colour = "#6D9EC1",
                                    size = 2, linetype = "solid"),
    panel.grid.major = element_line(size = 0.25, linetype = 'solid',
                                    colour = "black"), 
    panel.grid.minor = element_line(size = 0.25, linetype = 'solid',
                                    colour = "black")) + 
  theme(plot.background = element_rect(fill = "white"))  + 
  theme(strip.text = element_text(size = 10)) + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1, size = 10)) + 
  theme(axis.text.y = element_text(size = 10)) + 
  theme(axis.title.x = element_text(size = 10) + 
          theme(axis.title.y = element_text(size = 10), 
                theme(title =element_text(size = 10))))

#plot for Physics subject
physics_sector <- data.frame(
  Category = c("Astrophysics", "Condensed Matter", 
              "General Relativity and Quantum Cosmology", 
              "High Energy Physics", "Mathematical Physics",
              "Nonlinear Sciences", "Nuclear Physics" ,
              "Other Physics", "Quantum Physics"), 
  
  Amount = c(as.numeric(category_arxiv)[1], as.numeric(category_arxiv)[3],
             as.numeric(category_arxiv)[6] , as.numeric(category_arxiv)[7] ,
             as.numeric(category_arxiv)[9] , as.numeric(category_arxiv)[10],
             as.numeric(category_arxiv)[11], as.numeric(category_arxiv)[12],
             as.numeric(category_arxiv)[13]),
  
  stringsAsFactors = FALSE
)

plot2 <- ggplot(data = physics_sector) + 
  aes(Category, Amount) +   
  geom_bar(stat='identity', width=0.5, color='red',fill='midnightblue') +
  xlab("Physics Categories") + 
  ylab("Number of Papers Submitted") + 
  ggtitle("Number of Papers Submitted by Physics Category")  + 
  theme(panel.background = element_rect(fill = "pink1", colour = "midnightblue",
                                    size = 2, linetype = "solid"),
    panel.grid.major = element_line(size = 0.25, linetype = 'solid',
                                    colour = "black"), 
    panel.grid.minor = element_line(size = 0.25, linetype = 'solid',
                                    colour = "black")) + 
  theme(plot.background = element_rect(fill = "white"))  + 
  theme(strip.text = element_text(size = 10)) + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1, size = 10)) + 
  theme(axis.text.y = element_text(size = 10)) + 
  theme(axis.title.x = element_text(size = 10) + 
          theme(axis.title.y = element_text(size = 10), 
                theme(title =element_text(size = 10))))

ggarrange(plot1,plot2, nrow=2)

remove(category_arxiv)
remove(category_arxiv_final)

```

## Distribution Over Time
```{r,echo=FALSE}
#plot for all subjects
arxivsimplified = arxiv
arxivsimplified$Categories = str_replace_all(arxivsimplified$Categories, pattern="(Astrophysics)|(General Relativity and Quantum Cosmology)|(Condensed Matter)|(High Energy Physics)|(Mathematical Physics)|(Nonlinear Sciences)|(Nuclear Physics)|(Other Physics)|(Quantum Physics)" , "Physics")
arxivsimplified$Years = (as.numeric(format(datesarx,"%Y"))-1)%/%5*5

remove(datesarx)
```

```{r,echo=FALSE}
ggplot(data = arxivsimplified) + 
aes(x=Categories) + 
geom_bar( width=0.5, color='red',fill='steelblue', stat = "count") +
xlab("Categories") + 
ylab("Number of Papers Submitted") + 
ggtitle("Number of Papers Submitted by Category") + 
theme(panel.background = element_rect(fill = "#BFD5E3", colour = "#6D9EC1",
size = 2, linetype = "solid"),
panel.grid.major = element_line(size = 0.25, linetype = 'solid',
colour = "black"), 
panel.grid.minor = element_line(size = 0.25, linetype = 'solid',
colour = "black")) + 
theme(plot.background = element_rect(fill = "white")) + 
theme(strip.text = element_text(size = 10)) + 
theme(axis.text.x = element_text(angle = 20, hjust = 1, size = 8)) + 
theme(axis.text.y = element_text(size = 10)) + 
theme(axis.title.x = element_text(size = 10) + 
theme(axis.title.y = element_text(size = 10), 
theme(title =element_text(size = 10)))) + facet_wrap(~Years,scale="free", labeller=as_labeller(c(`1990` = "1991-1995",`1995` = "1996-2000",`2000`="2001-2005", `2005`="2006-2010", `2010`= "2011-2015", `2015`="2016-2020")))

remove(arxivsimplified)

```


## Active Institutions in Science
```{r,echo=FALSE}
authorsuni_2 <- as.data.frame(read.csv(file="./web-scraper/authorsuni.csv", stringsAsFactors=FALSE))
unicount <- table(authorsuni_2$University)

unicount_top10 <- as.data.frame(head(sort(unicount, decreasing=TRUE),n=11))[-1,]

ggplot(unicount_top10) + geom_col(aes(x=Var1,y=Freq), color = "red", fill = "midnightblue", width = 0.7) +
  xlab("Institution") + 
  ylab("Frequency") +
ggtitle("Most ArXiv Submitters by Institution")  + 
  theme(panel.background = element_rect(fill = "pink1", colour = "midnightblue",
                                    size = 2, linetype = "solid"),
    panel.grid.major = element_line(size = 0.25, linetype = 'solid',
                                    colour = "black"), 
    panel.grid.minor = element_line(size = 0.25, linetype = 'solid',
                                    colour = "black")) + 
  theme(plot.background = element_rect(fill = "white"))  + 
  theme(strip.text = element_text(size = 10)) + 
  theme(axis.text.x = element_text(angle = 50, hjust = 1, size = 10)) + 
  theme(axis.text.y = element_text(size = 10)) + 
  theme(axis.title.x = element_text(size = 10) + 
          theme(axis.title.y = element_text(size = 10), 
                theme(title =element_text(size = 10))))

remove(authorsuni_2)
```

## Statistical Data

```{r,echo=FALSE, fig.height= 10, fig.width=8}

coluni = read.csv(file = "Colleges_and_Universities.csv", stringsAsFactors = FALSE)

unicount = as.data.frame(unicount[-1], stringsAsFactors=FALSE)

unicount$Enrollment = coluni$TOT_ENROLL[match(table = coluni$NAME, x= toupper(as.character(unicount$Var1)))]

unicount$Size = coluni$INST_SIZE[match(table = coluni$NAME, x= toupper(as.character(unicount$Var1)))]

unicount$Employment = coluni$TOT_EMP[match(table = coluni$NAME, x= toupper(as.character(unicount$Var1)))]

sumsub = 1:5
for (i in 1:5){
sumsub[i] = sum(unicount$Freq[unicount$Size==i], na.rm=TRUE)
}
par(mfrow=c(2,1))
barplot(height = sumsub, main = "Total Submissions in Size Category", ylab = "Submission Count", col = "midnightblue")
plot(unicount$Size, unicount$Freq, ylab = "Submission Count", xlab = "Size of Institution", main = "Submissions by Size of U.S. Institution")





summary(lm(Freq ~ Size, data = unicount))


remove(unicount)
```

## Conclusion

- Graphs depicted are tip of the iceberg in terms of interesting trends
- Worldwide universities huge contributors of scientific journalism
- Physics originally dominated publishings, but seeing rapid growth in other fields

## Future Work

- Expand correlation dataset between authors and respective universities
  - Use GIS data to identify active research in world
  - Observe global statistics about authors
- Correlate university funding with publications to observe effect of wealth
- Identify growths in key words in papers over time to identify latest topics of interest

## Questions?
